---
output: word_document
---
Github repository: <https://github.com/bredowmax/matching>

### Load & optionally install required packages
- matchingR: Runs DAA and solves the Stable Marriage Problem by matching up agents in stable pairs
- dplyr:  Allows quick filtering in large datasets
- ggplot2: Renders graphs with a professional look
- rmarkdown: Export R code to Word
- ineq: Package that allows to display inequality
```{r, message = FALSE, warning = FALSE}
requiredPackages = c("matchingR", "dplyr", "ggplot2", "rmarkdown", "ineq", "reshape2", "ggthemes", "cowplot", "grid", "scales", "Rserve")
for(p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p)
  library(p,character.only = TRUE)
}
```

### Assumptions to be set by researcher
Sets network size, gender distribution, preference structure, experiment iterations and random distribution variables. Preference structures are defined as follows:

- *independent/subjective*: Each preference is generated randomly and has no correlation to any other preference. It can be phrased as: "Beauty lies in the eyes of the beholder"

- *correlated/objective/horizontal*: Perfectly correlated preferences indicate that all agents agree on a score. In this model, correlated preferenced are created through the average of all individually created preferences. A real life example is a unanimously valid beauty standard

- *intercorrelated/homolog/vertical*: Certain characteristics attract agents to each other, for example a low distance, a preferable age/height/height constellation, or common interests
```{r, message = FALSE, warning = FALSE}
nmen <- 5 #number of men
nwomen <- 3 #number of women
prefIndependent <- 1 #weight of subjective/independent preferences
prefCollective <- 0 #weight of collective/correlated preferences
prefHomolog <- 0 #weight of homolog/intercorrelated preferences
repetitions <- 0 #how many times the experiment is repeated with new set of random preferences
mn <- 6.3 #average preference
sd <- 1.9 #standard deviation
```

### Automatically set parameters
Only change if you know what you are doing
```{r, message = FALSE, warning = FALSE}
seed = 1 #sets seed for reproducability
seedMin=seed
seedMax=seed+repetitions
setSize=nmen+nwomen
sumPref=prefIndependent+prefCollective+prefHomolog
wI=prefIndependent/sumPref
wC=prefCollective/sumPref
wH=prefHomolog/sumPref
iterations=repetitions+1
percentileWomenFO=c()
percentileMenFO=c()
percentileWomenMO=c()
percentileMenMO=c()
```

### Defining Function to generate all preferences in one iteration
Each agent has independent, correlated or intercorrelated preferences for each agent of the other gender. The mix of these has been defined by the researcher in above code chunk. This model works with cardinal preferences, which allows not only to see who is preferred, but also by how much
```{r, message = FALSE, warning = FALSE}
generatePreferences <- function(nmen, nwomen, mn, sd, wI, wC, wH) {
  #generate individual cardinal preference matrices
  iUm <- matrix(rnorm(nmen*nwomen, mn, sd), nrow=nwomen, ncol=nmen) #independent pref. by men for women
  iUw <- matrix(rnorm(nwomen*nmen, mn, sd), nrow=nmen, ncol=nwomen) #independent pref. by women for men
  
  #Calculate attractiveness score of each agent
  Aw <- rowMeans(iUm) #average rating for each woman
  Am <- rowMeans(iUw) #average rating for each man
  
  #generate homolog preferences
  hUm <- matrix(rnorm(nmen*nwomen, mn, sd), nrow=nwomen, ncol=nmen) #homolog pref. by men for women
  hUw <- t(hUm) #homolog pref. by women for men (like each other exactly the same)
  
  #Calculate preferences stated by each agent
  uM <<- (wI*iUm + wC*Aw + wH*hUm) #weighted pref. by men for women
  uMt <<- t(uM)
  uW <<- (wI*iUw + wC*Am + wH*hUw) #weighted pref. by women for men
  uWt <<- t(uW)
  
  #sender
  senderMen = t(t(rep(1:nmen, each = nwomen)))
  senderWomen = t(t(rep(1:nwomen, each = nmen)))
  senderList = c(senderMen, senderWomen)
  
  #receiver
  receiverWomen = rep(1:nmen, times = nwomen)
  receiverMen = rep((nmen+1):(nmen+nwomen), times = nmen)
  receiverList = c(receiverWomen, receiverMen)
  
  #preferences
  preferenceList <- c(as.list(uM), as.list(uW))
  preferences = data.frame(cbind(senderList,receiverList,preferenceList))
  names(preferences) = c("swiperID", "targetID", "utility")
  return(preferences)
}
```

### Run Matching for i iterations
- Each experiment creates a new dating network
- Each network is two-sided and consists of n men and m women
- Each agent has random preferences for each agent of the other gender
- The preference structure is defined in the assumptions. Preferences can be independent, correlated or inter-correlated
  + The Deferred Acceptance Algorithm (DAA) solves the Stable Marriage Problem and matches up all agents in stable pairs
  + The DAA runs in two variants, which result in both the female-optimal (=male-pessimal) and the male-optimal (=female-pessimal) stable matching
  + If both results coincide, only one stable matching exists, which is optimal (and pessimal) for both genders
```{r, message = FALSE, warning = FALSE}
for (i in seedMin:seedMax){
  #set seed
  set.seed(i)
  
  #Generate Profiles
  preferences=generatePreferences(nmen, nwomen, mn, sd, wI, wC, wH)
  
  #generate Matchings
  resultsW <- galeShapley.marriageMarket(uW, uM) 
  resultsM <- galeShapley.marriageMarket(uM, uW) 

  #Record results for each agent in each iteration
  resultsWW <<- pnorm(diag(uW[resultsW$proposals,1:nwomen]),mean=mn,sd=sd)
  resultsWW[is.na(resultsWW)] <- 0 #this sets utility = 0 in case of no match
  resultsMM <<- pnorm(diag(uM[resultsM$proposals,1:nmen]),mean=mn,sd=sd)
  resultsMM[is.na(resultsMM)] <- 0
  resultsWM <<- pnorm(diag(uWt[resultsM$proposals,1:nmen]),mean=mn,sd=sd)
  resultsMW <<- pnorm(diag(uMt[resultsW$proposals,1:nwomen]),mean=mn,sd=sd)

  #Record Matching Satisfaction for all i iterations
  percentileWomenFO[[i]] = resultsWW
  percentileMenMO[[i]] = resultsMM
  percentileMenFO[[i]] = resultsMW
  percentileWomenMO[[i]] = resultsWM
}
```

###Generate Output
```{r, message = FALSE, warning = FALSE}
#Create columns separately
algorithm = matrix(c(rep("female-optimal", each=nwomen*iterations), rep("male-optimal", each=nmen*iterations)), nrow=(iterations*setSize), byrow=TRUE)
genderSender = matrix(c(rep("female", each = nwomen*iterations), rep("male", each = nmen*(repetitions+1))), nrow=(iterations*setSize), byrow=TRUE)
genderReceiver = matrix(c(rep("male", each = nwomen*iterations), rep("female", each = nmen*(repetitions+1))), nrow=(iterations*setSize), byrow=TRUE)
networkSize = matrix(setSize, nrow=iterations*setSize, byrow=TRUE)
weightMen = matrix(round(nmen/setSize, digits=2), nrow=iterations*setSize, byrow=TRUE)
weightWomen = matrix(round(nwomen/setSize, digits=2), nrow=iterations*setSize, byrow=TRUE)
weightIndependentPref = matrix(round(wI, digits=2), nrow=iterations*setSize, byrow=TRUE)
weightCollectivePref = matrix(round(wC, digits=2), nrow=iterations*setSize, byrow=TRUE)
weightHomologPref = matrix(round(wH, digits=2), nrow=iterations*setSize, byrow=TRUE)
matchSatisfactionSender = matrix(c(unlist(percentileWomenFO),unlist(percentileMenMO)), nrow=iterations*setSize, byrow=TRUE)
matchSatisfactionReceiver = matrix(c(unlist(percentileMenFO),unlist(percentileWomenMO)), nrow=iterations*setSize, byrow=TRUE)

#Combine to one dataset
tableau = cbind(algorithm, genderSender, genderReceiver, networkSize, weightMen, weightWomen, weightIndependentPref, weightCollectivePref, weightHomologPref, matchSatisfactionSender, matchSatisfactionReceiver)
colnames(tableau) = c("algorithm", "genderSender", "genderReceiver", "setSize", "weightMen", "weightWomen", "weightIndependentPref", "weightCollectivePref", "weightHomologPref", "matchSatisfactionSender", "matchSatisfactionReceiver")

#Export to Excel
filename=paste("sS_", nmen+nwomen, "_gD_", round(nmen/setSize*100), "_", round(wI, digits=2), "_", round(wC, digits=2), "_", round(wH, digits=2), ".csv", sep="")
write.csv(x=tableau, file=filename)
```
